% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{book}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\usepackage{graphicx}

\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Perfect Spider Cipher Theory},
  hidelinks,
  pdfcreator={LaTeX}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[paperwidth=5in,paperheight=8in,margin=0.5in]{geometry}
\usepackage[normalem]{ulem}
% Avoid problems with \sout in headers with hyperref
\pdfstringdefDisableCommands{\renewcommand{\sout}{}}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering

\title{\protect\hypertarget{perfect-spider-cipher-theory}{}{}Perfect
Spider Cipher Theory}
\author{}
\date{}

\begin{document}
\frontmatter
\maketitle

\mainmatter

This is a companion to \emph{Spider Cipher Theory} and uses its
notation: \(Z_{40}\) is the integers modulo \(40\), \(S_{40}\) the
arrangements of \(Z_{40}\), \(e = [ 0,1,...,39]\) the
sorted deck, \(t^{c}\) the cut at index \(c\), \(b_{i,j}\) the swap of
the cards at indexes \(i\) and \(j\), and {[}count down by odds, then up
by evens{]}

\(s = [ 39,37,35,33,31,...,9,7,5,3,1,0,2,4,6,8,...30,32,34,36,38]\)

the back-front shuffle. Composition is read right to left, so
\(s \circ t^{c}\) is a cut at \(c\) followed by a back-front shuffle.
Cycle notation \((a\ b\ c)\) means the permutation that sends \(a\) to
\(b\), \(b\) to \(c\), \(c\) to \(a\), and fixes everything else.

Facts written as FACTS(Name) are verified by the computer test suite in
\texttt{src/spider\_cipher\_facts.c}; each one re-checks on every build.

\hypertarget{the-perfect-shuffle}{%
\chapter{The Perfect Shuffle}\label{the-perfect-shuffle}}

The plain spider pseudo-shuffle has a short cycle:

\((s \circ t^{2})^{9} = e\)

Nine shuffles at cut index \(2\) return the deck exactly to where it
started, as demonstrated in FACTS(IReachable). Repeating a single cut,
the first return takes at least \(9\) steps, and \(9\) is achieved.

A sharper question: could some \emph{mixed} sequence of different cuts
close a loop in fewer than \(9\) shuffles? It cannot. A return of
length \(L\) is a closed walk \(e \rightarrow X \rightarrow e\); split
at its midpoint \(X\), reached by \(\lceil L/2 \rceil\) forward
shuffles and \(\lfloor L/2 \rfloor\) inverse shuffles. Any return of
length \(\leq 8\) would therefore place a non-identity deck in both the
radius-\(4\) forward ball and the radius-\(4\) inverse ball. Building
both balls (about \(2.6\) million decks each -- the full
\(40^{0} + \cdots + 40^{4}\) tree) and intersecting them by strict
sort, the only common deck is the identity. So no cut sequence returns
in fewer than \(9\) steps: the plain spider's first return is exactly
\(9\), over \emph{all} cut sequences. This meet-in-the-middle check
runs in a few seconds on a laptop (FACTS(ImperfectGirth)); it needs
only radius-\(4\) work because the midpoint halves the length.

Nine is far below what counting permits. The number of distinct
shuffle words of length at most \(30\) is
\((40^{31} - 1)/39 > 40! = \# S_{40}\), while for length at most
\(29\) it is less than \(40!\). So collisions among shuffle words can
not be forced before length \(30\), and a cycle floor of \(30\) is the
best that any family of \(40\) single-card-swap variants of this shape
could hope for.

The perfect spider shuffle adds one swap after the back-front shuffle:

\(q = b_{0,19} \circ s\)

The perfect spider graph has the same vertices \(V = S_{40}\) and, for
each vertex \(v\), the \(40\) outgoing edges
\(v \rightarrow q \circ t^{c} \circ v\).

With this one change the cycle floor rises from \(9\) to \(30\), one
step from the counting bound: for every cut index \(c\), the first
return of \((q \circ t^{c})^{k}\) to any cut of the starting deck takes
at least \(30\) steps. This is verified for all \(40\) cut indexes in
FACTS(Cycles), and the distinctness of all short shuffle words is
verified to depth \(6\) in FACTS(PerfectNeighborhood6).

\hypertarget{the-human-form}{%
\section{The Human Form}\label{the-human-form}}

Swapping the cards at indexes \(0\) and \(19\) after the shuffle sounds
like a job for a machine: nobody wants to count \(19\) cards into a
deck, twice, without error. But the same permutation has a second form:

\(b_{0,19} \circ s = s \circ b_{1,39}\)

To see this, read the placement off the shuffle \(s\): the back-front
shuffle delivers the card at index \(1\) (the second card) to index
\(19\), and the card at index \(39\) (the bottom card) to index \(0\).
So swapping indexes \(0\) and \(19\) after the shuffle is the same as
swapping indexes \(1\) and \(39\) before it. This identity is verified
on evolving decks for every cut index in FACTS(PerfectHumanForm).

The perfect shuffle is therefore the procedure:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  Cut the deck.
  \end{quote}
\item
  \begin{quote}
  Swap the second card with the bottom card.
  \end{quote}
\item
  \begin{quote}
  Back-front shuffle.
  \end{quote}
\end{enumerate}

No counting to \(19\) is ever required. The second card and the bottom
card are the two easiest cards to locate in a deck after the top card
itself.

\hypertarget{perfect-spider-graph-is-vertex-invariant}{%
\chapter{Perfect Spider Graph is Vertex
Invariant}\label{perfect-spider-graph-is-vertex-invariant}}

The proof is the same as for the plain spider graph. Every edge composes
a fixed permutation \(q \circ t^{c}\) on the left, so relabeling all the
vertices \(x\) by composing on the right, \(x \rightarrow x \circ (w^{- 1} \circ v)\),
makes vertex \(w\) look like vertex \(v\) and relabels all the edges
consistently. Nothing in the argument depends on which fixed
permutations the edges use.

As before, this means no amount of observing movement through the graph
reveals the starting point, and the structure of the graph can be
studied from any single vertex.

\hypertarget{perfect-spider-graph-is-fully-connected}{%
\chapter{Perfect Spider Graph is Fully
Connected}\label{perfect-spider-graph-is-fully-connected}}

Every vertex reaches every other vertex:
\(B_{\infty}(v) = S_{40}\). So restricting play to the reachable set
costs nothing; the active state space really is \(40!\).

The proof follows the same plan as for the plain spider graph, with
computer-verified waypoints. The plain proof leaned on the \(9\)-cycle
\((s \circ t^{2})^{9} = e\) at every turn -- the very cycle the perfect
shuffle exists to destroy -- so the perfect proof must fish with
different bait: it rides \emph{plain} cycles with \emph{perfect}
shuffles and keeps what washes back.

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  By vertex invariance we need only show \(B_{\infty}(e) = S_{40}\).
  A path \(e \rightarrow \cdot \cdot \cdot \rightarrow v\) means
  \(v = (q \circ t^{c_{p - 1}}) \circ \cdot \cdot \cdot \circ (q \circ t^{c_{0}})\)
  for some sequence of cuts, so each reachable \(v\) is shorthand for a
  sequence of cuts usable in any path.
  \end{quote}
\item
  \begin{quote}
  \(e\) and \(q^{- 1}\) are reachable. Choosing \(c = 0\) repeatedly
  reaches \(q^{k}\) for all \(k > 0\), and \(q\) has order \(36\):
  \(q^{36} = e\). So \(e\) is reachable as a cycle and
  \(q^{- 1} = q^{35}\) is reachable.
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches every cut: \(t^{a} = q^{35} \circ (q \circ t^{a})\).
  In words: take one step at cut \(a\), then \(35\) steps at cut
  \(0\); the shuffle-and-swap parts cancel and only the cut survives.
  Verified for all \(40\) cuts in FACTS(PerfectCuts).
  \end{quote}
\item
  \begin{quote}
  Trail fishing. For any swap \(b\) and any permutation \(w\),
  \(b \circ w = w \circ (w^{- 1} \circ b \circ w)\), and
  \(w^{- 1} \circ b \circ w\) is again a swap. So running the letters of
  a \emph{plain} cycle with \emph{perfect} shuffles pushes all the
  \(b_{0,19}\) factors through to one side, where they collect as a
  product of swaps -- the plain part vanishes entirely. The plain
  spider fact \((s \circ t^{39})^{24} = e\) (FACTS(InverseCycles))
  therefore lifts to a pure product of \(24\) swaps:

  \(f = (q \circ t^{39})^{24} = (4\ 38\ 17\ 19\ 29\ 30\ 35)\)

  a \(7\)-cycle, verified in FACTS(PerfectFish).
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches a \(3\)-cycle. The translate
  \(\bar{f} = t^{- 1} \circ f \circ t\) is reachable (cuts, then the
  fish, then cuts), and its support meets the support of \(f\) in
  exactly one card. Two cycles that share exactly one point have a
  \(3\)-cycle commutator:

  \(g = \bar{f}^{6} \circ f^{6} \circ \bar{f} \circ f = (30\ 35\ 31)\)

  using \(f^{6} = f^{- 1}\) since \(f\) is a \(7\)-cycle. Verified in
  FACTS(Perfect3Cycle).
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches \(a = (0\ 1\ 2)\). Conjugating by reachable elements
  keeps things reachable, and with \(H = t^{- 1} \circ q^{- 1} \circ t^{2} \circ q^{2}\),

  \(a = H \circ g \circ H^{- 1} = (0\ 1\ 2)\)

  which is a single composition anyone can verify by hand or machine.
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches every consecutive \(3\)-cycle
  \(a_{i} = t^{- i} \circ a \circ t^{i} = (i\ \ i \oplus 1\ \ i \oplus 2)\),
  where \(\oplus\) is \(+\) modulo \(40\).
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches the adjacent swap \(b_{0,1}\). The odd cut \(t\)
  composed with \(19\) consecutive \(3\)-cycles is exactly one swap:

  \(b_{0,1} = t \circ a_{2}^{- 1} \circ a_{4}^{- 1} \circ \cdot \cdot \cdot \circ a_{38}^{- 1}\)

  where \(a_{i}^{- 1} = a_{i}^{2}\) is reachable. (Equivalently:
  \(b_{0,1} \circ t^{- 1}\) is a \(39\)-cycle, and an odd-length cycle
  is a chain of \(3\)-cycles through every other point.)
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches every swap, exactly as in the plain proof:
  \(b_{i,i \oplus 1} = t^{- i} \circ b_{0,1} \circ t^{i}\); then
  \(b_{0,j} = b_{j - 1,j} \circ b_{0,j - 1} \circ b_{j - 1,j}\) by
  induction; then \(b_{i,j} = t^{- i} \circ b_{0,j - i} \circ t^{i}\).
  \end{quote}
\item
  \begin{quote}
  \(e\) reaches \(S_{40}\). Any arrangement can be sorted by swaps;
  applying the inverse swaps to \(e\) produces it.
  \end{quote}
\end{enumerate}

As an independent cross-check, a stabilizer-chain (Schreier--Sims)
computation of the group generated by
\(\{ q \circ t^{c}\}\) gives order exactly \(40!\).

\hypertarget{free-depth-and-the-birthday-bound}{%
\chapter{Free Depth and the Birthday
Bound}\label{free-depth-and-the-birthday-bound}}

Connectivity says every deck is reachable. The finer question, the one
that governs entropy capture, is how long the spider graph keeps
distinct paths distinct.

The \emph{free depth} \(r\) is the largest \(r\) such that all distinct
cut sequences of length at most \(r\) produce distinct decks. While the
tree is free, each step contributes \(\log_{2}40 \simeq 5.32\) bits of
path entropy faithfully; the first coincidence is the first time two
different histories become indistinguishable in the state.

Three quantities now separate. The \emph{cycle floor} is the shortest
first return of a single repeated cut. The \emph{girth} is the shortest
word of cuts equal to the identity. The free depth obeys
\(r \leq girth\), since an identity word of length \(L\) makes the word
of length \(L + 1\) that appends one more cut collide with that single
cut. For the plain spider, \((s \circ t^{2})^{9} = e\) forces
\(r \leq 9\). But nothing forces collisions to come from cycles at all.

\hypertarget{the-collapse-need-not-meet-at-the-identity}{%
\section{The Collapse Need Not Meet at the
Identity}\label{the-collapse-need-not-meet-at-the-identity}}

One would like to prove that the shortest collision factors through a
cycle -- that two colliding walks must, in effect, meet at the
identity -- because then the verified cycle floor would bound the free
depth and no neighborhood search would be needed.

It is false. In the \(10\)-card perfect model (the same construction
with \(Z_{10}\), swap of indexes \(0\) and \(4\)), the cut words

\(9,3,9,3\) and \(0,1,7,2,0\)

reach the same deck. This is a collision at depth \(5\), while the
girth of that graph is \(6\) and a cycle-induced collision needs
length \(girth + 1 = 7\). The collapse is a theta -- two paths
diverging and re-meeting away from the identity -- and no cycle
analysis can see it. This witness, and the exact floors, free depths,
and girths of the \(10\)- and \(12\)-card models, are direct
computations on those small decks -- intuition experiments, run by
hand rather than carried in the test suite.

\hypertarget{the-birthday-law}{%
\section{The Birthday Law}\label{the-birthday-law}}

What actually ends freeness is counting. The depth-\(r\) tree holds
\(40^{r}\) words mapping into \(40!\) decks, and coincidences among
them become unavoidable near \(40^{2r} \approx 40!\), that is,

\(r^{*} = \frac{1}{2}\log_{40}(40!) \approx 15.\)

The small models confirm that the free depth sits at this birthday
line no matter where the cycle floor is:

\begin{quote}
\begin{tabular}{rlrrr}
\(n\) & shuffle & floor & free depth \(r\) & \(\frac{1}{2}\log_{n}(n!)\) \\
\hline
10 & plain   & 4  & 3 & 3.3 \\
10 & perfect & 6  & 4 & 3.3 \\
12 & plain   & 6  & 5 & 4.0 \\
12 & perfect & 10 & 5 & 4.0 \\
14 & plain   & 6  & 5 & 4.8 \\
14 & perfect & 6  & 5 & 4.8 \\
\end{tabular}
\end{quote}

The floors wander from \(4\) to \(10\); the free depths track the last
column. (The observed first collisions even arrive somewhat later than
independent random sampling would predict -- the tree's shared
prefixes mildly suppress coincidences -- so the walk is no worse than
random, and slightly better.)

\hypertarget{consequences-for-the-design}{%
\section{Consequences for the
Design}\label{consequences-for-the-design}}

The birthday law assigns the cycle floor its true job: \emph{stay
above the birthday line so that cycles never bind before statistics
do}. For \(40\) cards the line is at about \(15\).

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  Plain spider has floor \(9 < 15\): it is cycle-bound, with
  \(r \leq 9\). Its own algebra cuts it down before counting does.
  \end{quote}
\item
  \begin{quote}
  Perfect spider has floor \(30 > 15\): it is birthday-bound, with
  expected free depth near \(15\) -- about \(80\) bits of faithfully
  captured path entropy -- and below that depth the expected collision
  mass \(40^{2r}/(2 \cdot 40!)\) is negligible.
  \end{quote}
\item
  \begin{quote}
  No shuffle of any design can exceed the birthday line. The perfect
  shuffle therefore gives up nothing: it converts the binding
  constraint from a repairable design flaw into a law of counting.
  \end{quote}
\end{enumerate}

What is certified for \(40\) cards: all words of length at most \(6\)
are distinct (FACTS(PerfectNeighborhood6), an exhaustive search of
about four billion decks), so \(r \geq 6\). Each further level costs a
factor of \(40\); certifying the expected \(r \approx 15\) by
exhaustion is out of reach, and the small-model law is the honest
evidence that the perfect spider walk behaves like a random one. The
collisions that eventually end freeness are statistical, not
structural -- which is exactly the property a cipher wants.

\hypertarget{the-entropy-capture-theorem}{%
\chapter{The Entropy Capture
Theorem}\label{the-entropy-capture-theorem}}

The birthday line says where the free depth probably sits; the free
depth \(r\) itself is a certified bound. This chapter states the
entropy theorem in terms of the bound. The birthday appears only as a
clearly-labeled estimate of where the bound likely extends.

Call an input sequence \emph{blind} if the cut choices
\(C = (c_{1},...,c_{k})\) are statistically independent of the deck
state -- the situation of random padding cards, which are drawn without
reference to the deck.

\hypertarget{capture-from-a-fixed-key}{%
\section{Capture from a Fixed Key}\label{capture-from-a-fixed-key}}

\textbf{Theorem.} In any vertex-invariant graph of out-degree \(40\)
with free depth \(r\): from any fixed starting deck, after
\(k \leq r\) blind steps the state entropy equals the input entropy
exactly,

\(H(X_{k}) = H(C).\)

With uniform cuts this is \(k\log_{2}40 \simeq 5.32k\) bits, and the
statement holds for every entropy measure at once (Shannon, collision,
min-entropy), because the proof is structural: free depth makes the
word map injective on words of length at most \(r\), vertex invariance
makes right-translation by the starting deck a bijection, and an
injective map preserves a distribution wholesale. Deterministic
(plaintext-driven) steps are bijections too, so entropy never
decreases at any depth; blind steps add their full measure while the
total word length stays within \(r\).

\hypertarget{cuts-are-far-away-and-also-adjacent}{%
\section{Cuts Are Far Away, and Also
Adjacent}\label{cuts-are-far-away-and-also-adjacent}}

The one obstruction to extending the theorem to arbitrary \emph{key
distributions} is instructive. First, a proposition that sounds like
safety: every nontrivial rotation is far away in the word metric.
If \(t^{\delta} = W\) for a word \(W\) of length \(j\), then by the
relabeling identity

\(g_{c} \circ t^{\delta} = q \circ t^{c + \delta} = g_{c + \delta}\)

the word \(W\) followed by the cut \(c\) collides with the single cut
\(c + \delta\), a tree collision at depth \(j + 1\). So free depth
\(r\) forces \(j \geq r\): \emph{no nontrivial rotation has a word of
length below \(r\)}.

But the same identity read the other way says rotations are adjacent
in the \emph{difference} metric:
\(t^{\delta} = g_{a}^{- 1} \circ g_{a + \delta}\). Consequence: two
keys differing by an initial cut, \(v\) and \(t^{\delta} \circ v\),
have identical one-step fans under a blind card --
\(\{ g_{c} \circ t^{\delta} \circ v\} = \{ g_{c + \delta} \circ v\}\)
as \(c\) ranges over all cuts -- so one blind card merges them
completely. Blindness cannot prevent this: a uniform cut shifted by
\(\delta\) is still uniform. No free-depth hypothesis can repair it,
because it is not a flaw: keys that are equal up to an initial cut
ought to be the same key.

\hypertarget{the-class-theorem}{%
\section{The Class Theorem}\label{the-class-theorem}}

So measure keys the way the merge demands: on \emph{cut classes},
decks up to an initial cut (\(40!/40 = 39!\) classes,
\(\log_{2}39! \simeq 153.8\) bits). Two facts make the class-level
account work.

First, classes branch at the full rate: the \(40\) children of any
deck lie in \(40\) distinct classes, because
\(q^{- 1} \circ t^{j} \circ q\) is never a rotation for
\(j \neq 0\) (a finite check, the depth-\(2\) case of the free
depth).

Second, class collisions reduce to deck collisions one level deeper:
if \(W = t^{j} \circ W'\) for words of length \(m\) and \(n\), then
appending the cut \(c\) to \(W\) and the cut \(c + j\) to \(W'\)
gives equal decks at lengths \(m + 1\) and \(n + 1\). So the class
free depth is at least \(r - 1\).

\textbf{Theorem.} For \emph{any} key distribution independent of \(k\)
uniform blind cards, with \(k\) at most the class free depth,

\(H(X_{k}) = H(class(X_{0})) + k\log_{2}40,\ \) exactly.

The proof absorbs the within-class offset of the key into the first
blind card by the relabeling identity -- a uniform first cut composed
with an unknown shift is a uniform first cut, independent of the
class -- and then applies the injectivity of the class-level word
map. For non-uniform blind cards the equality relaxes to
\(H(X_{k}) \geq H(class(X_{0})) + H(C)\).

Nothing is lost in passing to classes. The \(\log_{2}40 \simeq 5.32\)
bits of within-class position are not destroyed; they are
\emph{refreshed} by the first blind card, and the ledger balances
exactly:

\(\log_{2}40! = \log_{2}39! + \log_{2}40.\)

\hypertarget{bound-versus-estimate}{%
\section{Bound versus Estimate}\label{bound-versus-estimate}}

What the certificates give today: deck free depth \(r \geq 6\)
(FACTS(PerfectNeighborhood6)), hence class free depth at least \(5\),
hence the first \(5\) blind cards of the standard \(10\)-card prefix
are captured \emph{exactly} -- about \(26.6\) bits on top of the class
entropy of the key, by theorem. Each further certified level of \(r\)
converts another \(5.32\) bits from estimate to theorem.

What the estimate suggests: the birthday law places the free depth
near \(\frac{1}{2}\log_{40}40! \approx 15\), so the full \(10\)-card
prefix (\(53.2\) bits) is expected to be captured exactly, and beyond
the free depth the loss is governed by collision mass -- of order
\(40^{k}/40!\), under one bit all the way to saturation. But that is
an estimate. The theorem is the part that rests on \(r\), and \(r\)
is the part a machine can certify.

\hypertarget{why-not-swap-first-and-last}{%
\chapter{Why Not Swap First and
Last}\label{why-not-swap-first-and-last}}

There is a neighboring variant that looks just as good and is easier to
guess: swap the \emph{first and last} cards after the shuffle,

\(\hat{q} = b_{0,39} \circ s = s \circ b_{38,39}\)

which in human form is ``swap the bottom two cards, then shuffle'' --
one card away from the perfect shuffle's ``swap the second card with the
bottom card.'' It was implemented at one point in this project. It is
wrong, and this chapter records why, so it is not explored again.

\hypertarget{six-shuffles-return-the-deck}{%
\section{Six shuffles return the
deck}\label{six-shuffles-return-the-deck}}

\((\hat{q} \circ t^{4})^{6} = e\)

Six shuffles at cut index \(4\) return the deck \emph{exactly} to where
it started. The cycle floor is \(6\): worse than the plain spider's
\(9\), and a factor of five below the perfect shuffle's \(30\).
Verified in FACTS(FirstLastCycleFloor).

The floor matters because returning to (a cut of) the starting deck
repeats pad structure. The security story depends on the deck making
long excursions through \(S_{40}\) between any such coincidences; the
counting bound permits excursions of length \(30\), the perfect shuffle
achieves \(30\), and the first/last swap squanders the entire margin.

\hypertarget{nine-shuffles-are-one-swap}{%
\section{Nine shuffles are one
swap}\label{nine-shuffles-are-one-swap}}

\((\hat{q} \circ t^{2})^{9} = b_{1,39}\)

Nine shuffles at cut index \(2\) amount to swapping two cards and
changing nothing else, as verified in FACTS(FirstLastFish). A cipher
step is supposed to scramble; here nine of them barely stir the deck.

The mechanism is the trail fishing of the connectivity proof, working
against the variant. The plain pseudo-shuffle \(s \circ t^{2}\) has
orbits of sizes \(9,9,9,9,3\) with the fixed card \(38\), and its lone
\(3\)-orbit is \(\{ 0,1,39\}\). The trail of a swap rides
the orbits of its two endpoints. The first/last swap \(b_{0,39}\) has
\emph{both} endpoints inside the \(3\)-orbit, so its nine-swap trail
wraps a three-card track and collapses to a single swap. The perfect
swap \(b_{0,19}\) places only one endpoint there -- index \(19\) rides
a \(9\)-orbit -- so its trails spread across the deck instead of
collapsing.

\hypertarget{the-record}{%
\section{The Record}\label{the-record}}

The first/last graph is, for what it is worth, still fully connected:
its nine-shuffle swap \(b_{1,39}\) is itself the first rung of a
complete swap ladder, verified in FACTS(FirstLastCuts),
FACTS(FirstLastAdjacent), and FACTS(FirstLastSwaps). Connectivity is its
only virtue.

The lesson, recorded so this ground is not plowed twice: the cycle
floor is the gate. Any candidate variant of the spider shuffle must
first demonstrate its floor over all \(40\) cut indexes -- as
FACTS(Cycles) does for the perfect shuffle and
FACTS(FirstLastCycleFloor) does for this rejected one -- before any
other property earns attention. A variant that fails the gate fails,
no matter how natural the procedure feels in the hands.

\backmatter
\end{document}
